### 2014-10-14 R/allocation.Rnw
Problem description for Lu

\documentclass{article}

\usepackage{parskip}

\begin{document}

Dynamics of leaf mass ($M_F$) and woody biomass ($M_S$) depend on annual allocation of NPP to these pools ($\phi_F$ and $phi_S$), and turnover rates ($yr^{-1}$) ($r_F$ and $r_S$), as in:

\begin{equation}
\frac{dM_F}{dt} = \phi_F NPP - r_F M_F
\end{equation}

\begin{equation}
\frac{dM_S}{dt} = \phi_S NPP - r_S M_S
\end{equation}
With the constraint that $\phi_F + \phi_S = 1$.

We can further assume that NPP is proportional to foliage biomass,

\begin{equation}
NPP = a_F M_F
\end{equation}
So that Eq. 1 and 2 become,

\begin{equation}
\frac{dM_F}{dt} = (\phi_F  a_F - r_F) M_F
\end{equation}

\begin{equation}
\frac{dM_S}{dt} = \phi_S  a_F  M_F - r_S M_S
\end{equation}

I don't know how to deal with this system of DE's, what I would like to have is a solution that gives me the dependence of $M_F$ on $M_S$ (and time of course),

$$M_F = f(M_S,t)$$

Or if it is more convenient,

$$M_F/M_S = f(M_S,t)$$

where either function of course also depends on the various parameters. If it helps the solution, you may also assume that $r_F >> r_S$

\end{document}