### 2014-10-17 manuscript/draftpaper.Rnw
Major figures update. Also added species-level plot. Manuscript results section improved.

\documentclass[a4paper]{article}

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
\bibliographystyle{plain}

\begin{document}

\title{Specific leaf mass explains difference in biomass partitioning between woody plant functional types}

\author{Remko Duursma, Daniel Falster and BAAD contributors}

\maketitle

% Notes from Daniel phone call
% - make case that allocating biomass vs. targeting certain leaf area vs. stem area is very different
% 
% fig 1 : more grey levels. scale for grey levels
% figure captions need to be stronger statements
% 
% use all data with leaf mass, explain that subset has same results
% 
% Fig 9 + 10 instead of Fig 2 in main text.
% Then, Fig 3 says that SLM explains differences.
% 
% Fig 3 + 6 log-log axes
% 
% Fig 4 first, introduce PFT differences
% 
% No difference in m.so/a.astba2, but large difference m.lf/a.stba2
% 
% remove fig 5
% 
% combine fig 3
% 
% climate-related effects are captured by shift in PFT
%

%--------------------------------------------------------------------------------------------%

<<preload, echo=FALSE>>=
library(knitr)
opts_chunk$set(echo=FALSE)

load("tables/Tables_varpart.RData")
baad <- readRDS("../data/baad.rds")
@

%--------------------------------------------------------------------------------------------%

\section{Introduction}

\begin{enumerate}
\item The importance of biomass allocation in global vegetation models \cite{ise_comparison_2010}, feedback on fluxes of CO2, lifetime of C in the system \cite{friend_carbon_2014}, which depends on partitioning of biomass into short-lived foliage vs. long-lived woody pools.

\item There is no shortage of theoretical developments in understanding biomass allocation (\cite{cannell_carbon_1994,franklin_modeling_2012,enquist_land_2012}). In particular, theoretical work focusing on optimal partitioning theory is promising (e.g. \cite{valentine_modeling_2012}) but is yet to lead to operational models for dynamic global vegetation models.

\item Previous syntheses on plant biomass relationships has predominantly focused on the universal nature of size-biomass relationships in plants. Little is known about differences between plant functional types (evergreen vs. deciduous, angiosperm vs. gymnosperm). Since most global models parameterize allocation submodels by PFT, it is imperative that we better understand these potential differences.

\item Major plant functional types differ widely in specific leaf mass (SLM, the ratio of leaf mass to area, an index of leaf thickness) (\cite{poorter_causes_2009}), which can be expected to affect biomass partitioning in several ways. First, SLM correlates with leaf lifespan. Longer lived leaves would lead to higher plant leaf mass ($M_F$) as long as annual production rates are similar (\cite{enquist_global_2002}). Another interpretation is that plants may target a certain leaf area display given a certain biomass investment in supporting tissues, because light interception and photosynthesis correlate with the exchange area of the plant.

\item It has been reported numerous times that gymnosperms (with higher SLM) have higher leaf mass fraction than angiosperms (with lower SLM) \cite{enquist_global_2002, mccarthy_consistency_2007, poorter_biomass_2012, sardans_tree_2013}, although a recent study reported no difference (\cite{wolf_allometric_2010}). No study to date has been able to analyze differences between deciduous and evergreen angiosperms, or link differences in partitioning between PFTs to other plant traits such as SLM.

\item What is lacking is a solid empirical base that can be used to constrain any model of allocation, that includes woody plants of all sizes, of all major plant functional types and biomes. Most global efforts to date use the Cannell database of forest biomass, which is not based on individual biomass estimates. Here we use a new database (Falster et al.), the largest ever compiled for individual woody plants, including data for major plant functional types (PFTs), and all major biomes.

\item We consider here the relationship between leaf biomass and area with stem diameter (an important allometric relationship, Enquist others, and related to the pipe model if the ratio sapwood / heartwood is relatively constant, an important measure of hydraulic supply), the aboveground leaf mass fraction (partioning of biomass into leaf vs. woody pools), and the leaf area ratio (a measure of the photosynthetic area per unit woody biomass). We expect that $M_F$ and $A_F$ are approximately proportional to the square of stem diameter, which would remove size-related shifts in allometry that often confound interpretations. For each of these variables, we ask whether 1) there are differences between major plant functional types after removing size-related variation, 2) there are differences due to climate, compared between major biomes, or as a function of the mean annual temperature (as recently reported for whole stand biomass, \cite{reich_temperature_2014}).

\end{enumerate}

%------------------------------------------------------------------------------------%
\section{Methods}

\begin{itemize}
\item BAAD database (Falster et al., in review), including a total of \Sexpr{nrow(baad$data)} observations, but here we use only data that include leaf mass ($M_F$), leaf area ($A_F$), aboveground woody biomass ($M_W$), total root biomass ($M_R$), plant height ($H$), and stem area measured at ground level ($A_S$), or at breast height (ca. 1.3m) ($A_{Sbh}$).  For each analysis, we therefore used different subsets of the data (far fewer observations are available for root biomass, in particular), sample sizes are included in the caption of each figure (\emph{to be added}). We further excluded deciduous gymnosperms (since very few data were available), and glasshouse and common garden studies, including only field grown woody plants (including plantations).

\item We focus here only on aboveground biomass because we found no PFT differences in root-shoot scaling (Fig.~\ref{fig:figureSI6}), and approximately isometric scaling (sma slope = xxx) (consistent with other reports, \cite{hui_near_2014, cairns_root_1997}).

\item For the aboveground biomass pools, we calculated the leaf mass fraction ($M_F / M_T$, where $M_T$ is total aboveground biomass) and leaf area ratio ($A_F / M_T$). LMF, LAR, SLM, $M_F$ and $A_F$ are related by,

$$LMF = M_F / M_T = (M_F / A_F) (A_F / M_T) = SLM * LAR$$

\item We can further decompose aboveground biomass ($M_T$) into a size-related metric (basal stem area, $A_S$), and the ratio of $M_T$ to $A_S$, which is a function of wood density and plant woody volume to basal area.

$$LMF = M_F / M_T = (M_F / A_S) (M_T / A_S)^{-1}$$

We test for PFT differences to both these components of biomass partioning. We were unable to further split $A_S/M_W$ because wood density data were available only for a subset of data. Chave has shown that etc. etc.

\item We considered three plant functional types : evergreen angiosperms, evergreen gymnosperms, and decidous angiosperms. All locations were further separated into boreal (including sub-boreal), temperate, and tropical biomes (decided by contributing author, but checked with MAT and latitude).

\item We did not attempt here to include SLA from another database (such as TRY), instead only used SLA directly estimated for the harvested plants.

\item Stem area was measured either at breast height (ca. 1.3m) or at the base of the plant (normally for small plants). For the subset of the data where both were measured, we estimated basal stem area (Asba) from breast height area (Asbh) by fitting XXX equation. Results were similar when we used either measure of stem area (not shown).

\item We used a generalized additive model (GAM) to visualize the relationships, but these were not used for formal inference.

\item We used linear mixed-effects models (with species within study as the random effect) to test for significance and to calculate confidence intervals for PFT-specific averages. Least-square means for LMF and LAR were calculated with the 'lsmeans' package.

\item All variables were log-transformed prior to analysis, to stabilize variance.

\item Variance partitioning with marginal (fixed effects) and conditional (fixed and random effects) $R^2$ for linear mixed-effects models as reported by \cite{nakagawa_general_2013}.

\item For conifers, leaf area was converted to half-total surface area, which is most appropriate in terms of light interception  \cite{lang_application_1991, chen_defining_1992}, for comparison to flat leaves.
\end{itemize}

%------------------------------------------------------------------------------------%
\section{Results}

\begin{itemize}

\item We found large differences between PFTs in biomass partitioning between leaves and stems (Fig.~\ref{fig:figure2}, see also Fig.~\ref{fig:figureSI3}). Calculated least-square means of LMF and LAR (averages accounting for variation explained by plant height) showed large differences in mass-based allocation between PFTs (LMF), and small inconsistent differences in area-based allocation (LAR) (Fig.~\ref{fig:figure2}, see Fig.~\ref{fig:figureSI4} for raw data on leaf area ratio by PFT).

\item The ratio of aboveground biomass to basal stem area did not differ by PFT (Fig.~\ref{fig:figureSI5}).

\item However we found very large differences between PFTs in the ratio of leaf mass to basal stem area (Fig.~\ref{fig:figure3}), but the ratio leaf area to basal stem area converged between PFTs.

\item We then expressed the average $M_F/A_s$ as a function of the specific leaf mass (SLM = $M_F / A_F$), a measure of the leaf construction cost to the plant. We found that, across PFTs, $M_F/A_S$ was approximately proportional to the average specific leaf mass for the PFT. As a result of this approximate proportionality, the ratio of leaf area to basal stem area ($A_F/A_S$) was not different between PFTs (Fig.~\ref{fig:figure4}) (see also Fig.~\ref{fig:figureSI6} for raw data).

\item This relationship also held approximately across species (Fig.~\ref{fig:figureSI7}).

\item Because there was substantial variation between individual plants (Figs.~\ref{fig:figure3}), we estimated the variance explained by PFT, plant height, and biome in $M_F/A_S$ using a linear mixed-effects model (see Table~\ref{varpartpipe}). We found that plant height and biome explained only \Sexpr{100*signif(Table_pipemodel_varpart$Marginal[2],2)}\%, but when PFT was added to the model, the variance explained increased to \Sexpr{100*signif(Table_pipemodel_varpart$Marginal[3],2)}\%. If we used SLM as a predictor instead of PFT, the $R^2$ was similar (\Sexpr{100*signif(Table_pipemodel_varpart$Marginal[4],2)}\%), showing that PFT affected $M_F/A_S$ largely through differences in SLM. Consistent with this, PFT, plant height and biome explained very little variance in $A_F/A_S$ (Table~\ref{varpartpipe}).

\item We also partionined the variance of LMF and LAR (Table~\ref{varpartpipe}). When added to a model that included plant height and biome, PFT increased the variance explained in log-transformed LMF from \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[2],2)} to \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[3],2)}\%. However, SLM was not a good substitute for PFT, as the explained variance decreased to \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[4],2)}\%. For (log-transformed) LAR, PFT and SLM explained very little additional variance.

\item Although biome was a significant predictor for all variables tested, it consistently explained very little variation (Table~\ref{varpartpipe} and~\ref{varpartlmflar}). Pair-wise comparisons between biomes within PFT mostly did not show significant differences for $M_F/A_S$ or $A_F/A_S$. The only exception was a higher $A_F/A_S$ for tropical compared to temperate evergreen angiosperms (Fig.~\ref{fig:figureSI8}).

\item Unlike \cite{reich_temperature_2014}, we found no effect of mean annual temperature on the scaling of leaf biomass with woody biomass (Fig.~\ref{fig:figureSI9}).

\end{itemize}

%------------------------------------------------------------------------------------%
\section{Discussion}

\begin{itemize}

\item We found coordination between an important leaf trait and whole-plant biomass partitioning. This has implications for global vegetation models.

\item How is allocation implemented in current DGVMs? Either constant allocation fraction, target allometric equations, or optimization \cite{de2014does}. Constant allocation fraction gives constant LAR by pft if leaf lifespan is exactly proportional to SLM. Target allometric equations can be made pft-invariant by using allometry of leaf area, not leaf mass. Optimization models probably already give the right answer because normally LAI is optimized (Woodward).

\item A recent study compared allocation routines in a number of leading ecosystem models \cite{de2014does}, and recommended to constrain allocation by observed biomass fractions instead of using constant allocation fractions, which is common practice.

\item As a first approximation, which is useful for large scale DGVMs, models should aim to keep leaf area to stem cross-sectional area constant, and vary biomass allocation accordingly. Specific leaf area then gives the ratio of leaf biomass to stem biomass.

\item We found no effects of climate on anything. This is different from \cite{reich_temperature_2014} but they studied whole-stand biomass, not individual plant biomass. So perhaps this means the climate effects are on stand density not individual plant effects. No effect of climate is good news for models that parameterize the globe by PFT, as this seems sufficient, based on our dataset. Are there experimental data that show temperature effects on allocation? What about water availability?

\end{itemize}

%------------------------------------------------------------------------------------%
\section{Lit notes}

Reich et al 1992, for seedlings, negative relationship leaf lifespan and LAR. This is different from our results, since SLM is nearly proportional to leaf lifespan, and we found no relationship between LAR and SLM (but do they have all the PFTs??). However, across stands, positive relationship between leaf biomass and leaf lifespan, which is consistent with our results.

Cornelissen et al 1996 positive relationship SLA and LAR, but this is for tiny seedlings harvested 21 days after first leaf appearance. But check references and who cites this!

\cite{friend_carbon_2014} showed that 30\% of the uncertainty in global vegetation carbon simulated with seven vegetation models is due to the residence time of carbon in vegetation. This residence time depends on turnover rates of the various plant pools, and hence also on the distribution of biomass between these pools. For example, vegetation with a higher leaf mass fraction will have a lower overall residence time of carbon.

\cite{de2014does} "We have shown that allocation approaches that are constrained by biomass fractions (such as functional relationships) were more successful at capturing observed trends, and were generally more robust, than approaches based on allocation coefficients."

\cite{friedlingstein_toward_1999} make no mention of plant functional type differences or the need to vary allocation by pft.

\cite{givnish_adaptive_2002} discusses costs and benefits of evergreen vs. deciduous habit, in terms of constructing a growth model that takes into account photosynthesis, leaf lifespan, belowground costs. However he does not consider or report on differences in allocation associated with leaf lifespan. Instead he simply states that 'leaf allocation ratio' relates to leaf lifespan, which of course is only true if total leaf mass is constant.

\cite{callaway_biomass_1994} example of lower leaf mass / sapwood area for desert population of PIPO vs. montane population. For pointing out small-scale climatic effects on allometry (that may not be obvious in global scale analyses).

\cite{poorter_biomass_2012} found a slightly larger LMF in gymnosperms, but the difference was small and could not split between deciduous and evergreen angiosperms due to a lack of data. They argue that "most conifers retain leaves for two or more years. Thus, the higher LMF may simply be the consequence of leaf longevity rather than higher assimilate partitioning per se. In this way, conifers are able to compensate for their lower physiological activity per gram of leaf." This raises the question whether LMF differs by angiosperms and gymnosperms, or as a consequence of leaf longevity (especially deciduous vs. evergreen).

\cite{enquist_global_2002} report 2.6 times higher leaf mass at given stem mass for conifers compared to angiosperms, but did not separate further, nor interpreted this difference in terms of SLA. Their interpretation is 'that conifers typically retain three cohorts of leaves that have less well-developed [...] mesophyll'. Not sure how that latter bit makes sense, and leaf lifespan interpretation is no good; because it still begs the question why would conifers not simply have 1/3 less leaf mass per cohort?

\cite{mccarthy_consistency_2007} further demonstrate large difference in LMF between angiosperms and conifers, and note that a large chunk of the residuals is correlated with SLA and LL. They interpret this rather vaguely as 'leaf morphology may constrain leaf mass distribution'. Their SLA was not on the same individuals, but they could have calculated PFT average LAR anyway, but did not.

\cite{oneill_comparative_1981} reported much higher leaf biomass for gymnosperms vs. angiosperms, at a given woody biomass increment. This means (as Niklas reported) a much higher productivity per unit leaf biomass for angiosperms.

Serious lack of biomass data to constrain models. Studies like \cite{wolf_forest_2011} resort to studying individual tree allometry based on dividing stand biomass by stand density. \cite{wolf_allometric_2010} did not find any differences in biomass fractions (root,stem,leaf) between angiosperms and gymnosperms (but again this was based on stand biomass data, but still this is strange!).

\cite{hui_near_2014} Isometric scaling of above vs. belowground biomass in chinese forest biomass database.

\cite{sardans_tree_2013} report similiar LMF for evergreen angio. and gymnosperms, and ca. 40\% less leaf mass for decid. angio. Local study, 3000 plots in Catalonia.

\cite{reich_temperature_2014} "On average, when standardized for MAT and Mstem, conifers tended to have more Mfol and less Mroot than angiosperm forests". Their Fig. 2 shows trends in LMF with MAT etc., in bins of stem size. Also allometric analysis with MAT as covariate, showing significance of MAT. For Angiosperms, our explanation would be that evergreen angio is more common at high MAT. But what about conifers?

\cite{cairns_root_1997} found no difference in root-shoot ratio between gymno and angio.

\cite{delucia_climate-driven_2000} LA/SA decreases with summer VPD for Pinus but not non-Pinus gymno. Also Pinus has much lower LA/SA than non-Pinus; suggesting that we should separate gymnosperms into those groups as well.

\cite{grier_notes:_1974} early study on leaf mass proportionality to sapwood area (and uses mass not area).

\section{Tables}

<<echo=FALSE, message=FALSE, warning=FALSE>>=
library(xtable)
@

<<results='asis'>>=
print(xtable(Table_pipemodel_varpart, label="varpartpipe", 
             caption="Explained variance in the ratio of leaf mass to basal stem area, and leaf area to basal stem area. The marginal R2 is that explained by the fixed effects only (H = plant height, B = biome, PFT = plant functional type. The conditional R2 is that explained by fixed and random effects (species within study). All fixed effects were highly significant (P \\textless 0.001). In addition to the fixed effects shown, all interactions were added to each of the models."), 
      include.rownames=FALSE, caption.placement="top")
@

<<results='asis'>>=
load("tables/Tables_varpart.RData")
print(xtable(Table_LMFLAR_varpart, label="varpartlmflar",
      caption="Explained variance in the leaf mass fraction (leaf mass / aboveground biomass) and leaf area ratio (leaf area / aboveground biomass). See Table~\\ref{varpartpipe} for more explanation. For plant height, we also added the squared term as a fixed effect, and the interaction with PFT or SLM, because the relationship was non-linear."),
      include.rownames=FALSE, caption.placement="top")
@

\clearpage
\section{Figures}

\begin{figure}[h!]
    \centering
    \includegraphics{Figure1_MAPMAT_baad_vs_worldclim.pdf}
    \caption{Mean annual precipitation and mean annual temperature for all included studies. Hexagons are terrestrial observations extracted from Worldclim (darker bins indicating more observations). Note that Worldclim observations include land areas where woody vegetation does not occur (desert, tundra). The deciduous gymnosperms were not included in any analysis in this paper, but are shown here.  }
    \label{fig:figure1}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure2_LMF_lines_lsmeans_3panel.pdf}
    \caption{Leaf mass fraction (leaf mass / aboveground biomass) by plant functional type. Each symbol is an individual plant. Lines are generalized additive model fits.Bottom panels show average leaf mass fraction (LMF) and leaf area ratio (LAR) accounting for variation explained by plant height (least-square means) with a mixed effects model (with species within study as a random effect). Error bars are 95\% confidence intervals.}
    \label{fig:figure2}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure3_hist_alfast_mlfast.pdf}
    \caption{Plant functional types diverge strongly in mass-based allocation, but converge in area-based allocation. Shown are histograms of leaf mass per unit basal stem area ($M_F/A_S$) and leaf area per unit basal stem area ($A_F/M_S$) grouped by the three major plant functional types. Vertical lines and shading are means with confidence intervals (same as in Fig.~\ref{fig:figure2}). Histograms of $M_F/A_S$ show large differences between PFTs, while $A_F/A_S$ converges to a single average across all PFTs, though much variation exists between plants.}
    \label{fig:figure3}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure4_mlf_alf_astbaest_pftmeans.pdf}
    \caption{Average leaf mass per unit basal stem area (panel 1), and leaf area per unit basal stem area (panel 2) for the three major plant functional types. Error bars are 95\% confidence interval for the mean (calculated with a mixed-effects model with species within study as the random effect). Letters denote significant differences. }
    \label{fig:figure4}
\end{figure}

\clearpage
\section{Supporting Information Figures}

\renewcommand\thefigure{S\arabic{figure}}    
\setcounter{figure}{0}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureSI-1_MAPMAT_vegetation.pdf}
    \caption{Mean annual precipitation and temperature for all contributed datasets, by vegetation type. Note that in this study, we did not include glasshouse studies.}
    \label{fig:figureSI1}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureSI-2_mrt_mso_bypft.pdf}
    \caption{Near-isometric scaling of above-ground with below-ground biomass, and no effects of plant functional type. The black line is a 1-1 line.}
    \label{fig:figureSI2}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureSI-3_mlfmst_byht_pft.pdf}
    \caption{Leaf and woody biomass scaling with plant height.}
    \label{fig:figureSI3}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureSI-4_LAR_pft_lines.pdf}
    \caption{Leaf area ratio (leaf area / aboveground biomass) by plant functional type. Each symbol is an individual plant. Lines are generalized additive model fits.}
    \label{fig:figureSI4}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureSI-5_mso_astba2_twopanel.pdf}
    \caption{Relationship between aboveground biomass and basal stem area, each point is an individual plant. Lines are spline fits. The right panel shows least-square means of the ratio of aboveground biomass to basal stem area, estimated from a linear mixed-effects model with basal stem area (linear and squared terms) and PFT. }
    \label{fig:figureSI5}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureSI-6_mlf_alf_astba2_bypft.pdf}
    \caption{Plant leaf mass as a function of basal stem area (left panel), and leaf area as a function of basal stem area (right panel). Each symbol is an individual plant. Lines are generalized additive model fits.}
  \label{fig:figureSI6}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureSI-7_mlf_alf_ast_byspecies.pdf}
    \caption{Plant leaf mass as a function of basal stem area (left panel), and leaf area as a function of basal stem area (right panel). Individual plant sata were averaged by species/study (the same species occuring in multiple datasets thus appears multiple times). Lines are linear regressions with 95\% confidence intervals. Both regressions were significant, but R2 was 0.28 in the left panel, and only 0.02 in the right panel. The coefficient of variation between species was 0.19 for log-transformed MF/AS, and only 0.099 for AF/AS.}
  \label{fig:figureSI7}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureSI-8_mlf_alf_astbaest_pftlongmeans.pdf}
    \caption{Like Fig.~\ref{fig:figure2}, but for five combinations of major plant functional type and biome. Out of 9 possible biome / pft combinations, only five in the dataset had leaf area estimates (and two combo's are not in the dataset, e.g. tropical evergreen gymno). Note that within pft, vegetation is never significant (this result will also be in another figure!).}
    \label{fig:figureSI8}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureSI-9_MAT_LMFscaling.pdf}
    \caption{For each species within study, we fit the scaling relationship of leaf mass as a function of woody biomass assuming an exponent of 3/4, using standardized major axis. The resulting coefficient (b0) is size-invariant (and thus superior to the simple ratio of leaf mass to woody biomass, which is strongly size-dependent, see Fig.~\ref{fig:figure3}). We found no effects of mean annual temperature on the coefficient. The triangles are from Roth et al. (2007), a study on Pinus eliottii with exceptionally high leaf mass per unit woody biomass.}
  \label{fig:figureSI9}
\end{figure}

\clearpage
\section{Supporting Information Tables}

<<tablepreload>>=
load("tables/Table_counts.RData")
printit <- function(x, caption){
  print(xtable(x, align="rlll|l",
               caption=caption
               ), hline.after=c(0,3), caption.placement="top")

}
@

<<results='asis'>>=
printit(tab_mlfastba, caption="Number of observations used in the analysis of $M_F/A_S$ and $A_F/A_S$. Individual plants (number of species). DA = Deciduous Angiosperm, EA = Evergreen Angiosperm, EG = Evergreen Gymnosperm.")
@

% <<results='asis'>>=
% printit(tab_alfastba, caption="Number of observations used in the analysis of Af/As. Individual plants (number of species). DA = Deciduous Angiosperm, EA = Evergreen Angiosperm, EG = Evergreen Gymnosperm.")
% @
%

<<results='asis'>>=
printit(tab_lmf, caption="Number of observations used in the analysis of leaf mass fraction (LMF) and leaf area ratio (LAR). Individual plants (number of species). DA = Deciduous Angiosperm, EA = Evergreen Angiosperm, EG = Evergreen Gymnosperm.")
@

% 
% <<results='asis'>>=
% printit(tab_lar, caption="Number of observations used in the analysis of leaf area ratio (LAR). Individual plants (number of species). DA = Deciduous Angiosperm, EA = Evergreen Angiosperm, EG = Evergreen Gymnosperm.")
% @

\clearpage
\bibliography{allometry}

\end{document}